Evaluation of Machine Translation in Languages

### 2024-07-06 paper/paper.tex
updated

Machine translation (MT) has emerged as a pivotal technology in facilitating cross-cultural communication and global information exchange. With the exponential growth of digital content in diverse languages, the demand for accurate and efficient translation systems has never been greater. This paper aims to evaluate and compare the performance of various machine translation models across multiple languages, shedding light on their effectiveness and limitations in real-world applications.

Effective machine translation hinges on the ability of algorithms to comprehend and accurately render meaning from source text in one language to target text in another. Traditionally, statistical methods dominated the landscape, employing linguistic rules and alignment techniques. However, the advent of neural machine translation (NMT) has revolutionized the field, leveraging deep learning architectures to achieve remarkable improvements in translation quality.

Statistical machine translation, which previously dominated MT research for many years with its reliance on various count-based models, has largely been surpassed by neural machine translation (NMT) \cite{stahlberg-2020-nmt-review}.

The languages chosen for evaluation in this study represent a wide overview, striking a balance between resources and diversity.

The evaluation criteria encompass standard metrics such as BLEU (Bilingual Evaluation Understudy) \cite{papieni-2002-bleu}, METEOR (Metric for Evaluation of Translation with Explicit ORdering) \cite{lavie-2007-meteor} to capture nuances in translation quality. These metrics not only quantify the fidelity of translations but also offer insights into the models' adaptability and robustness across different linguistic pairs.

@@

Evaluating languages enhances understanding of linguistic structures, phonetics, syntax, semantics, and other linguistic components. This knowledge contributes to the broader field of linguistics and helps in developing theories about language evolution, language families, and the cognitive processes involved in language acquisition and use.

Machine translation can be tracked to as early as 1949 \cite{weaver-1999}

@@

mBART \cite{liu-2020-mbart} is considered one of the State-of-the-Art as of this moment for multilingual translation.

T5 \cite{raffel-2023-t5}

\section{Tatoeba Dataset}

@@

@@

\section{Methodology}

Take the list of most spoken language by population according to Wikipedia \cite{wikipedia-list-languages}, then take one for each branch.

Sentences in English are downloaded, 1,898,494 sentences (it is unclear why it is less than the number stated in the Tatoeba website). Then for each langauges, download sentence pairs compared to English. Merge every language sentences into one big dataframe, only keep where sentences exist for every language.

Target language will be English, with machine translation task defined as Many-to-English.

>>>>

Machine translation (MT) can be traced down to as early as 1949 \cite{weaver-1999}, recently growing as a popular research field within Natural Language Processing (NLP).

Statistical machine translation, which previously dominated MT research for many years with its reliance on various count-based models, has largely been surpassed by neural machine translation (NMT) \cite{stahlberg-2020-nmt-review}. Nowadays, Large Language Models (LLMs) such as mBart \cite{liu-2020-mbart}, M2M-100 \cite{fan-2020-m2m100}, NLLB-200 \cite{nllb200}, T5 \cite{raffel-2023-t5}, and GPT-4 \cite{openai-2024-gpt4-technical-report} dominated the field of machine translation. DeepL \cite{deepl} currently proclaimed itself as 'The world most accurate translator'.

@@

@@

\section{Dataset}

@@

To build the dataset,

Sentences in English are downloaded, 1,898,494 sentences (it is unclear why it is less than the number stated in the Tatoeba website). Then for each langauges, download sentence pairs compared to English. Merge every language sentences into one big dataframe, only keep where sentences exist for every language

@@

\section{Methodology}.

The languages chosen for evaluation in this study represent a wide overview, striking a balance between resources and diversity. Each language will then be translated into English (Many-to-English) and evaluated. The evaluation criteria encompass standard metrics such as BLEU (Bilingual Evaluation Understudy) \cite{papieni-2002-bleu}, METEOR (Metric for Evaluation of Translation with Explicit ORdering) \cite{lavie-2007-meteor} to capture nuances in translation quality. These metrics not only quantify the fidelity of translations but also offer insights into the models' adaptability and robustness across different linguistic pairs.
